Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-10 vesicle_code/docs/adhesion/adhesion.tex
touched up figure 11

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2.
  \label{eqn:SLP}
\end{align}
\todo[inline]{Define $\rr$ and other variables in
equation~\eqref{eqn:SLP}}
The traction jump of the velocity $\uu(\xx) = \SS[\ff](\xx)$ is the
density function $\ff$, so the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{qua-bir2016}.  The SDC method is able
to achieve second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

@@

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti et
al.~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significant increases the intrinsic viscosity at all reduced areas.  To
further characterize the effect of adhesion, in the right plot of
Figure~\ref{fig:shearIntrinsicViscosity} we decompose the intrinsic
viscosity into the contributions from the bending and tension, and the
contribution from the adhesion.  We see that the effect of the adhesion
on the intrinsic viscosity is largest for vesicles with small reduced
areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\sigma = 0.7$,
  and the adhesion length scale is $\delta = 0.4$.  The length of the
  vesicles at each of the reduced areas is fixed, while the area changes
  with differing reduced areas.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti et
  al.~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em Right}:
  The decomposition of the intrinsic viscosity into the contributions
  for the bending and tension of the vesicles, and the contribution from
  the vesicle adhesion.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral Equation Formulation}
Because the fluid equations are linear and elliptic, an integral
equation formulation is possible.  This has several advantages,
including that only the vesicle interfaces require discretization.  We
start by defining the Stokes single-layer potential
\begin{align}
  \SS[\ff](\xx) &= \frac{1}{4\pi\mu} \int_\gamma \left(
    -\log \rho + \frac{\rr \otimes \rr}{\rho^2} \right) 
    \ff(\yy) ds_\yy, \quad \xx \in \RR^2,
  \label{eqn:SLP}
\end{align}
where $\rr = \xx - \yy$ and $\rho = \|\rr\|$.  The traction jump of the
velocity $\uu(\xx) = \SS[\ff](\xx)$ is the density function $\ff$, so
the fluid velocity is $\xx \in \RR^2$ is
\begin{align*}
  \uu(\xx) = \uu_{\infty}(\xx) + \SS[\xxi](\xx).
\end{align*}
The traction jump is given by three terms: the bending, tension, and
adhesion.  Therefore
\begin{align*}
  \xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA \xx.
\end{align*}
Applying the no-slip boundary condition on each vesicle, the governing
equations are
\begin{align*}
  &\dot{\xx} = \uu_{\infty}(\xx) + \SS[\xxi](\xx), \\
  &\xx_s \cdot \uu_s = 0, \\
  &\xxi = -\kappa_b \xx_{ssss} + (\sigma \xx_s)_s + \AA\xx.
\end{align*}
Defining the bending operator as $\BB[\xx](\ff) = -\kappa_b \ff_{ssss}$,
and the tension operator $\TT[\xx](\sigma) = (\sigma \xx_s)_s$, the
no-slip boundary condition can be written as
\begin{align*}
  \pderiv{\xx}{t} = \SS \BB \xx + \SS \TT \sigma + \SS \AA \xx.
\end{align*}
We apply the IMEX-Euler method 
\begin{align*}
  \frac{\xx^{N+1} - \xx^N}{\Delta t} = \SS^N \BB^N \xx^{N+1} + 
  \SS^N \TT^N \sigma^{N+1} + \SS^N \AA^N \xx^N,
\end{align*}
that treats both the bending and tension terms
semi-implicitly~\cite{qua-bir2014}. Note that the adhesive term is
treated explicitly.  Therefore, a single time step requires solving
\begin{align}
  \xx^{N+1} - \Delta t \SS^N \BB^N \xx^{N+1} - 
    \Delta t \SS^N \TT^N \sigma^{N+1} = \xx^N + 
    \Delta t \SS^N \AA^N \xx^N,
  \label{eqn:IMEXEuler}
\end{align}
along with the inextensibility constraint that is discretized as
\begin{align*}
  \xx_s^{N} \cdot \xx_{s}^{N+1} = 1.
\end{align*}
We discretize the vesicles at a set of collocation points, compute the
bending and tension terms with Fourier differentiation, and apply
Alpert quadrature to the logarithmically singular single-layer potential
$\SS$~\cite{alp1999}.  The source and target points of the adhesion
force never coincide since they are always on different vesicles, so the
adhesion force~\eqref{eqn:adhesionForce} is computed with the spectrally
accurate trapezoid rule~\cite{tre-wei2014}.  To control the error and
achieve higher-order accuracy in time, we apply a time adaptive spectral
deferred correction method~\cite{qua-bir2016}.  The SDC method is able
to achieve second-order accuracy by iteratively applying the first-order
method~\eqref{eqn:IMEXEuler}.

@@

In the left plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare
the intrinsic viscosity of a single tank treading vesicle to a doublet.
To validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti et
al.~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behaves similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\sigma = 0.7$,
  and the adhesion length scale is $\delta = 0.4$.  The length of the
  vesicles at each of the reduced areas is fixed, while the area changes
  with differing reduced areas.  The black marks denote intrinsic
  viscosity values computed by Ghigliotti et
  al.~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em Right}:
  The decomposition of the intrinsic viscosity into the contributions
  for the bending and tension (blue) of the vesicles, and the
  contribution from the vesicle adhesion (red).  Also included is twice
  the intrinsic viscosity of a single tank-treading vesicle.}
\end{figure}
